---
title: trailcountR
output: github_document
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

## Package Description

<!-- badges: start -->
<!-- badges: end -->

The goal of trailcountR is to make processing raw trail counter data easier. Land
managers may use a variety of methods to gather visitor traffic counts
at the trail or forest level. These methods may include deploying infrared counters
at trail heads (TRAFx), magnetic vehicle counters, in-person parking lot counts, 
cameras and others. These raw formats with instantaneous or hourly counts must be
transformed into daily counts and combined to be used in downstream analyses.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("OutdoorRD/trailcountR")
```
## Example 1

Once you have parsed IR data for a specific device, the days included will only 
be dates that the device logged a count. If the site had no visitors for a day 
or the device was broken, then no counts are logged for that day and thus that 
date will not be included in the data. To create a complete timeseries for IR 
data, use the `make_complete_timeseries()` function. Any dates where the counter
was not triggered will have an NA count. 
```{r example1}
library(trailcountR)
library(magrittr)

# create a complete timeseries for IR data
df <- tibble::tibble(
	PlacementID = c(102, 102, 102, 102),
	date = c("19-06-04", "19-06-07", "19-06-08", "19-06-10"),
	time = c("14:00", "14:00", "13:01", "12:02"),
	count = c(15, 13, 12, 50))
df %>% 
	kableExtra::kbl() %>% 
	kableExtra::kable_material_dark(lightable_options = "striped")
```

```{r make_complete_timeseries}
# make the complete timeseries
df2 <- make_complete_timeseries(df)
df2 %>% 
	kableExtra::kbl() %>% 
	kableExtra::kable_material_dark(lightable_options = "striped")
```
## Example 2
Once a complete timeseries has been generated for an IR counter, there might be 
certain days which have missing counts. This could occur when no visitors
triggered the counter for that day or when the counter has stopped working. For
downstream analyses, these missing counts should be replaced with a count of zero. 
The `insert_zero_counts()` function makes this happen. 

```{r example2}
library(trailcountR)
library(magrittr)

## Make toy IR counter data
df <- tibble::tibble(
	PlacementID = c(102, 102, 102, 102),
	date = c("19-06-04", "19-06-07", "19-06-08", "19-06-10"),
	time = c("14:00", "14:00", "13:01", "12:02"),
	count = c(15, 13, 12, 50))

# complete the timeseries
df2 <- make_complete_timeseries(df)

df2 %>% 
	kableExtra::kbl() %>% 
	kableExtra::kable_material_dark(lightable_options = "striped")
```

```{r insert_zero_counts}
# replace the missing counts with zeros 
df3 <- insert_zero_counts(df2)
df3 %>% 
	kableExtra::kbl() %>% 
	kableExtra::kable_material_dark(lightable_options = "striped")

```

